---
title: "Seed Rain Analysis"
author: K. Carter Wynne
output: 
  html_document:
    smart: no
    theme: flatly
    float: yes
    css: "style.css"    
editor_options: 
  chunk_output_type: console
mainfont: Arial
mathfont: Arial
fontsize: 12pt
---

#  {.tabset .tabset-pills .tabset-fade}

## Setup and Overview

### **Quantifying seed rain patterns in a remnant and a chronosequence of restored tallgrass prairies in north central Missouri**

**Background**: Seed rain, or the dispersal of seeds into an area, is an understudied phenomenon closely related to the establishment and persistence of plant species in a community. However, natural seed rain patterns in tallgrass prairie communities and how these patterns change during the restoration process remain poorly understood. To address these knowledge gaps, we revisited a seminal grassland seed rain study conducted in 1980 at a remnant Missouri tallgrass prairie and quantified seed rain patterns in the same remnant prairie and a chronosequence of nearby restorations. In May of 2019, we deployed 200 artificial turf grass seed traps (0.1 x 0.1 m) in 2, 6-7, and 15-year-old restored and remnant north-central Missouri tallgrass prairies. We collected and replaced seed traps every 2-weeks from June to December 2019.

**Author:** Katherine Wynne ([Wynnekat\@msu.edu](mailto:Wynnekat@msu.edu){.email})

**Co-authors:** M. Parker-Smith, E. Eyerly, L. Sullivan

**Created:** 24 January 2023

------------------------------------------------------------------------

**Files:**

1)  *Seed_Rain_Data_2019_FINAL.xlsx* - contains seed rain data

2)  *Seed_Rain_Date.xlsx* - contains lumped seed data per week for each site

------------------------------------------------------------------------

**R Version:** R 4.2.2

**RStudio Version:** 2023.03.0+386

**Package Version:** Found in the .readme document in GitHub Repository

**Last updated: 10 April 2023**

------------------------------------------------------------------------

### Libraries and Files

```{r, eval=TRUE, echo=TRUE, warning=TRUE, message=FALSE}

## Clear environment 
rm(list=ls())

## Libraries
# Import excel files

library(readxl)

# Data cleaning and visualization
library(tidyverse)

# Community ecology functions (NMDS, diversity indices, etc.)
library(vegan)

# Package needed to do contrasts
library(emmeans)

# Export dataframes to excel files
library(writexl)

# Dimensions check for NMDS stress
library(goeveg)

# Needed in order to install and load the pairwiseAdonis package
## install.packages("devtools")
## library(devtools)
## install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")

# Function for pairwise PERMANOVA tests
library(pairwiseAdonis)

# Function that allows us to fit a GLM with the negative binomial distribution
library(MASS)

# Makes grided figures
library(cowplot)

# Functions that let you plot ordiellipse in ggplot
library(tcltk)
library(BiodiversityR)


# Function that repels geom_text so that text labels don't overlap
library(ggrepel)

# Function that helps me create tables
library(flextable)

# Function that allows me to export tables to word
library(officer)

# Function that gets means, sds, etc. easily
library(mosaic)
```

```{r, eval=TRUE, echo=FALSE, warning=TRUE, message=FALSE}

# Data sets

Seed_Cleaning <- read_excel("Seed_Rain_Data_2019_FINAL.xlsx")
Seed_Rain_Date <- read_excel("Seed_Rain_Date.xlsx")
Species_data <- read_excel("Species_list_SR_detailed.xlsx")
```

```{r, results = 'hide',echo=FALSE,  message = FALSE, warning = FALSE}

## Data Cleaning
### General Data Cleaning 


# Checking to make sure all the SPP6 levels are correct (no misspellings or weirdness)

sort(unique(Seed_Cleaning$SPP6))
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}

## Code to remove non-viable seeds from the dataset
Seed_Data <- filter(Seed_Cleaning, !grepl('NONE',SPP6)) 
```

```{r, results = 'hide',echo=FALSE, message = FALSE, warning = FALSE}

# Code to lump certain species together
Seed_Data_Lumped <- Seed_Data
for(i in 1:nrow(Seed_Data)){
  
  if(Seed_Data_Lumped[i,5] == "AGAFAS"){Seed_Data_Lumped[i,5] <- "AGASPP"}
  if(Seed_Data_Lumped[i,5] == "EUPSER"){Seed_Data_Lumped[i,5] <- "EUPSPP"}
  if(Seed_Data_Lumped[i,5] == "GENPUB"){Seed_Data_Lumped[i,5] <- "GENSPP"}
}  

### Species remaining after lumping and removing unknowns
list(sort(unique(Seed_Data_Lumped$SPP6)))
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}

# Make sure variables are in the correct format
Seed_Data_Lumped$Plot <- as.factor(Seed_Data_Lumped$Plot)
Seed_Data_Lumped$Number <- as.numeric(Seed_Data_Lumped$Number)
Seed_Data_Lumped$Week <- as.factor(Seed_Data_Lumped$Week)
Seed_Data_Lumped$Transect <- as.factor(Seed_Data_Lumped$Transect)
Seed_Data_Lumped$SPP6 <- as.factor(Seed_Data_Lumped$SPP6)


# Remove columns unnecessary for analysis (Sampling Date, Notes, Taxon, Revisit)


# Data set with sampling date (analysis 2 and 4)
Seed_Data_Date <-  Seed_Data_Lumped[, -c(7,8,9)]


# Data set without sampling date (analysis 1 and 3)
Seed_Data_Lumped <- Seed_Data_Lumped[, -c(3,7,8,9)]
  
```

```{r}

# Get rid of unknown species to make a species list for analysis 4
Species_list_SR <- filter(Seed_Data_Lumped, !grepl('UNK',SPP6)) 
Species_list_SR <- filter(Species_list_SR , !grepl('unk',SPP6)) 
Species_list_SR <- filter(Species_list_SR , !grepl('	
PANICUM?',SPP6)) 


# Make a species list dataset
Species_list_SR <- Species_list_SR %>% 
  group_by(SPP6) %>% 
  summarize(totseeds = sum(Number)) 
Species_list_SR <- Species_list_SR[,-2]


#write_xlsx(Species_list_SR, "Species_list_SR.xlsx")
```

## Analysis 1: Are there differences in the quantity of seed rain between sites?

### **Are there differences in the quantity of seed rain between sites?**

To test for differences between the number of seeds falling in remnant and restored prairies, we fit a generalized linear model predicting number of seeds captured as a function of site (n = 4). We used a negative binomial distribution ("MASS" package; Venables & Ripley, 2002) to account for overdispersion present in the data. We then conducted a posthoc analysis using Dunnett-style contrasts ("emmeans" package; Lenth, 2023) with a multiple comparison adjustment for three tests to determine whether restored sites significantly differed in the number of seeds falling compared to the remnant prairie.

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

### Analysis 1: Data Wrangling


# Make a new dataframe (Uni_Seed) to analyze whether there are differences in the amount of seed dispersal occurring between sites
Uni_Seed <- Seed_Data_Lumped %>% 
  
# Make new column named "UniqueTran" to make a unique ID for each transect at each site 
## (e.g. Tucker Prairie Transect 1 = TP_1)
  
  unite("UniqueTran", Plot:Transect, sep = "_", remove = FALSE) %>% 
# Sum the total of seeds caught for each unique transect 
## (totseeds = total number of seeds captured in each transect)
  
  group_by(Plot, Transect, UniqueTran)  %>% 
  summarize(totseeds = sum(Number))


# Make Tucker Prairie (the remnant prairie) the reference level
Uni_Seed$Plot <- factor(Uni_Seed$Plot, levels=c("TP", "PFCA 1", "PFCA 2", "PFCA 3"))
str(Uni_Seed$Plot)
```

### Analysis 1: Fitting the Model

The mean and variance are not equal and indicates extreme overdispersion, meaning we cannot use a Poisson distribution.

```{r, message = FALSE, echo = FALSE, warning = FALSE}

# Checking for over dispersion
Uni_Seed %>% 
  group_by(Plot) %>% 
  summarise(Mean = mean(totseeds),
            Var = var(totseeds))
```

Below is a negative binomial model predicting the number of seeds captured as a function of site.

```{r, message = FALSE, echo = FALSE, warning = FALSE}

# Fit a negative binomial model predicting number of seeds by site
neg_binom_seeds_mod <- MASS::glm.nb(totseeds~Plot, data=Uni_Seed)

# Convert back to glm() style object  (look into this again you may not need to do it)
NegBinom_seeds_glm <- MASS::glm.convert(neg_binom_seeds_mod)

# Summary results of model
### Dispersion = 1 since k was already estimated (see John Snyder's notes)
summary(NegBinom_seeds_glm, dispersion = 1)
```

Posthoc comparisons between the remnant and reconstructed prairies.

```{r, message = FALSE, echo = FALSE, warning = FALSE}

## Do restored sites differ from the remnant?


# Contrasts
standardize <- emmeans(NegBinom_seeds_glm, "Plot")
contrast(standardize, "trt.vs.ctrl")



## 95% confidence intervals 

emmeans::emmeans(NegBinom_seeds_glm, "Plot", type = "response")

## Make a new dataframe to hold the model estimates and 95% confidence interval bounds

response <- c(2081, 6240, 2408, 1388)
lower <- c(1439, 4316, 1665, 959)
upper <- c(3009, 9023, 3482, 2007)
site <- c("TP", "PFCA 1", "PFCA 2", "PFCA 3")

neg.binom.est <- data.frame(site, response, lower, upper)

```

### Analysis 1: Visualizing Differences

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

# Make Tucker Prairie (the remnant prairie) the reference level

Uni_Seed$Plot <- factor(Uni_Seed$Plot, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))

neg.binom.est$site <- factor(neg.binom.est$site, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))
```

```{r, message = FALSE, echo = FALSE, results = 'hide', echo = FALSE, warning = FALSE}

# scatterplot with 95% confidence intervals

# setwd("~/Desktop/Seed Rain 2019/Graphics")
abund.jitter.plot <- ggplot()+
  geom_jitter(data=Uni_Seed, aes(x=Plot, y=totseeds, color = Plot), show.legend = FALSE, width = .2, size = 2) +
    geom_point(data = neg.binom.est, aes(x=site, y=response), show.legend = FALSE, size = 3.5)+
  geom_errorbar(data = neg.binom.est, aes(x= site, y = response, ymin = lower, ymax = upper), width = 0.15, cex = 0.75)+
  theme_classic()+
  xlab("Site")+
  ylim(0,10000)+
  ylab("Number of seeds")+
  scale_x_discrete(labels = c ("PFCA 1" = "Young", "PFCA 2" = "Middle", "PFCA 3" = "Old", "TP" = "Remnant"), guide = guide_axis(angle = 0)) +
  scale_color_manual(name = "Site", values=c( "#1f78b4", "#a6cee3", "#b2df8a","#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant"))+
   scale_x_discrete(labels = c ("PFCA 1" = "Young", "PFCA 2" = "Middle", "PFCA 3" = "Old", "TP" = "Remnant"), guide = guide_axis(angle = 0)) +
 theme(text=element_text(size=16), legend.key.size=unit(1, "cm"))+ geom_vline(xintercept = 3.5, linetype = "longdash")


abund.jitter.plot + annotate("text", x = 1 , y = 9500, label = "*", size = 9.5)
```

**Figure 1**: The number of seeds captured per transect in the remnant prairie (dark green), young (2-year-old, dark blue), middle-aged (6-7-year-old, light blue), and old (15-year-old, light green) reconstructed prairies between May 31st and December 12th, 2019. Error bars represent 95% confidence intervals around model estimates (black). Asterisk symbols indicate sites that had significantly different amounts of seeds captured when compared to the remnant prairie (p \< 0.05).

### Analysis 1: Results

The quantity of seed rain in reconstructions decreased with age. By at least 6-7 years post-initial seeding, the number of seeds dispersing in reconstructions was comparable to the remnant prairie in 2019. Only the youngest reconstructed prairie had significantly more seeds dispersing than the remnant prairie sampled in the same year.

## Analysis 2: Are there differences in the timing of seed rain between sites?

### **Are there differences in the timing of seed rain between sites?**

Similar to Rabinowitz and Rapp (1980), we graphed seed rain phenology to compare seed rain timing between restoration ages, site types, and study years

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}


### Analysis 2: Data Wrangling


# Let's do seed rain density to compare to Rabinowitz and Rapp 1980
Seed_Rain_Date$Plot <- as.factor(Seed_Rain_Date$Plot)

# ----------

# Make a new dataset for the PFCA and TP 2019 Sites 
Density_Date_PFCA_TP19 <- Seed_Rain_Date %>% 
  filter(Plot != "TP_1980")

# Make a new column for seed rain density per week for PFCA and TP 2019


## Density = height of carpet trap * width of carpet trap * # of carpets at a site
## Density = 0.1m * 0.1m * 50
## Density = 0.5m^2 (for a site)


Density_Date_PFCA_TP19$Density_J <- Density_Date_PFCA_TP19$Number_J / 0.5

# ----------

# Make a new dataset for just TP in 1978


## Density = pi * radius of trap^2 * # of petri dishes at TP
## Density = 3.145 * 0.045m^2 * 50
## Density = 0.31807688 m^2 


Density_Date_TP1980 <- Seed_Rain_Date %>% 
  filter(Plot == "TP_1980")


Density_Date_TP1980$Density_J <- Density_Date_TP1980$Number_J /  0.31807688
New_Seed_Rain_Date <- full_join(Density_Date_TP1980, Density_Date_PFCA_TP19)
```

### Analysis 2: Visualizing differences

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

# TP 2019, TP 1978, PFCA 1, PFCA 2, PFCA 3

# png("Seed_abundance_overtime_R78_SRP.png", width=850, height =600)
ggplot(New_Seed_Rain_Date, aes(y=Density_J, x=Date, group=Plot)) +
  geom_line(aes(color=Plot), linewidth = 1.5, alpha = 1) +
  geom_point(aes(shape=Plot, fill=Plot), size=3, alpha = 1)+
  scale_color_manual(name = "", values=c("#1f78b4", "#7ab1cf", "#8bc454", "#288022", "#3A3B3C"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP_2019", "TP_1980"), labels = c("Young", "Middle", "Old", "Remnant 2019", "Remnant 1978"))+
  scale_shape_manual(name = "", values=c(23,24,22,21, 20),breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP_2019", "TP_1980"), labels = c("Young", "Middle", "Old", "Remnant 2019", "Remnant 1978")) +
  scale_fill_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c", "#3A3B3C"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP_2019", "TP_1980"), labels = c("Young", "Middle", "Old", "Remnant 2019", "Remnant 1978"))+
  theme_classic()+
  scale_x_datetime(date_breaks = '1 month', date_labels="%b %d", limits=as.POSIXct(c("2019-06-5", "2019-12-25")))+
  theme(text=element_text(size=16), legend.key.size=unit(0.25, "cm"), legend.position="bottom", legend.box = "horizontal")+
  ylab("Seeds"~m^-2)+
  xlab("Date")+
  guides(x = guide_axis(angle = 45))+
  ylim(0, 32000)
```

**Figure 2.** Timing of seed rain in reconstructed prairies of various ages and a remnant prairie. Data for the remnant in 1978 obtained from Rabinowitz and Rapp (1980).

### Analysis 2: Results:

Seed rain in all prairies exhibited a bimodal pattern of dispersal, where seed rain density peaked once in the early summer and again in the fall. However, compared to 1978, the timing of peak seed rain density was shifted two weeks earlier in 2019. Timing of the seed rain converged with the remnant prairie (2019) as the prairie reconstructions matured. As expected, the youngest (2-year-old) reconstructed prairie was the most divergent, where the timing of the first peak in seed rain was delayed by a month compared to the other prairies sampled in 2019. Overall, the timing of the seed rain captured in the middle-aged (6-7-year-old) and old (15-year-old) reconstructed prairie closely resembled the timing of the remnant prairie in 2019.

## Analysis 3: Are there differences in the composition of seed rain between sites?

### **Are there differences in the composition of seed rain between sites?**

To assess for seed rain compositional differences between sites, we first created a community distance matrix using Bray-Curtis dissimilarity, which considers changes in species abundance between sites. Before calculating Bray-Curtis dissimilarity, we relativized species abundance by site total (across plot rows). We then visualized compositional differences with non-metric multidimensional scaling (NMDS) ordination. As Clarke (1993) recommended, we did not use solutions with stress \> 0.20. We used the envfit function (permutations = 5000) found in the "vegan" package (Oksanen et al., 2020) to identify species significantly associated with driving the observed compositional differences among sites and plot them as vectors (p \< 0.001).

Using permutational multivariate analysis of variance (PERMANOVA, permutations = 999) and posthoc pairwise comparison tests ("pairwiseAdonis" package; Martinez Arbizu, 2020), we examined potential differences in species composition between remnant and restored prairies. Additionally, we tested for homogeneity of multivariate dispersions to assess whether there were differences in variation among communities between study sites. We first calculated multivariate dispersion for each site using the *betadisper* function, and then conducted an ANOVA to detect differences between sites. We used the "vegan" package (Oksanen et al., 2020) to conduct all multivariate analyses.

*Note that using Jaccard dissimilarity and presence/absence produced similar results to Bray-Curtis (Not shown).*

```{r, results = 'hide',echo = FALSE, message = FALSE, warning = FALSE}


### Analysis 3: Data Wrangling
Seed_Com_Long <- Seed_Data_Lumped %>% 
  
# Make new column named "UniqueTran" to make a unique ID for each transect at each site
## (e.g. Tucker Prairie Transect 1 = TP_1)
  
  unite("UniqueTran", Plot:Transect, sep = "_", remove = FALSE) %>% 
# Sum the total of seeds caught for each species at a unique transect 
## Need to do this because I changed the label for certain species (e.g., SIMPIL -> SYMSPP)
  
  group_by(Plot, Transect, UniqueTran, SPP6)  %>% 
  summarize(totseeds = sum(Number))
# Reorder dataset to "wide" format where each SPP6 factor gets its own column and the value is according to number
  
seed_community_wide_full <- spread(Seed_Com_Long, SPP6, totseeds, fill = 0)


# ----------

# Remove all columns that isn't a species ID or Site

seed_community_wide <- seed_community_wide_full[, -c(1,2)]

# Make row labels UniqueTran
seed_community_wide_labs   <- seed_community_wide[,-1]
row.names(seed_community_wide_labs) <- seed_community_wide$UniqueTran


# Convert into matrix
seed_mat <- data.matrix(seed_community_wide_labs)
```

### Analysis 3: NMDS

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

# Standardize the community matrix
## By site total

seed_mat_tot <- decostand(seed_mat, "total")
```

```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE, fig.show='hide'}

# Check the number of dimensions needed for the NMDS  (write the result of this test)

## Bray-Curtis dissimilarity
dimcheckMDS(seed_mat_tot, distance = "bray", k=6, trymax= 50, autotransform = FALSE)


###
## *** Best solution repeated 5 times
# 0.34302456 0.17850813 0.11694087 0.08649123 0.06448893 0.05274240
# We can use 2 dimensions (< 20)
```

```{r, echo = FALSE, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

Tot_NMDS <- metaMDS(seed_mat_tot, dist = "bray", k=2, trymax=1000, autotransform =  FALSE)
```

```{r, results = 'hide', message = FALSE, echo = FALSE, warning = FALSE}

### Looks at which species were the most influential in driving the observed patterns

inner.spp.fit <- envfit(Tot_NMDS, seed_mat_tot, permutations = 5000, k =2)

#head(inner.spp.fit)
```

```{r, results = 'hide', message = FALSE, echo = FALSE, warning = FALSE, fig.show='hide'}

# Make a grouping variable so I can match site names to the NMDS output
site = c(rep("PFCA 1", 10), rep("PFCA 2", 10), rep("PFCA 3", 10), rep("TP", 10))

# Put the NMDS scores into it's own dataframe 
scrs <- as.data.frame(scores(Tot_NMDS, display = "sites"))

# Match the NMDS scores to the right site labels
scrs <- cbind(scrs, site)

# Extracts the required scaling factor for the vector arrows
arrow_factor <- ordiArrowMul(inner.spp.fit)

# Take the envfit model and convert into a a dataframe (scaled)
spp.scrs <- as.data.frame(scores(inner.spp.fit, display = "vectors"))*arrow_factor

# Add the p values, R squared values to the envit model species score dataframe
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs), Pvalues = inner.spp.fit$vectors$pvals, R_squared = inner.spp.fit$vectors$r)


# Subset out species that had a p value <= 0.001 (since I think env fit truncates to 0.001)
spp.scrs.01 <- subset(spp.scrs, Pvalues < 0.002)
```

```{r, results = 'hide', message = FALSE, echo = FALSE, warning = FALSE, fig.show='hide'}

# Only including this plot because I need to call the plot in order to obtain the ordiellipse object

plot(Tot_NMDS$points, xlim = c(-2, 2), ylim = c(-2, 2), col = alpha(0.5), xlab = "NMDS1", ylab = "NMDS2", cex.lab = 1.25, cex.axis = 1.25, bty="n")


# Store the ordiellipse data into a list
site.ellipses <- ordiellipse(Tot_NMDS, groups = site, display = "sites", kind = "se", conf = 0.95)


# Convert the ordiellipse data into a dataframe that I can plot in ggplot
site.ellipses.long2 <- ordiellipse.long(site.ellipses, grouping.name="site")
```

```{r, results = 'hide', message = FALSE, echo = FALSE, warning = FALSE}

# Plot the species scores + the NMDS results

NMDS_plot <- ggplot(scrs) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, color = site, fill = site, shape = site), size=2.5, alpha = 0.8) + 
  scale_shape_manual(name = "", values=c(23,24,22,21),breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019")) +
  scale_color_manual(name = "", values=c("#1f78b4", "#7ab1cf", "#8bc454", "#288022"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019"))  +
  scale_fill_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019")) +
  geom_polygon(data = site.ellipses.long2, aes(x = axis1, y=axis2, colour = site, fill =   after_scale(alpha(colour, 0.2))), size = 0.2, show.legend = FALSE) +
  coord_fixed() +
  geom_segment(data = spp.scrs.01, 
             aes(x = 0, xend = NMDS1 * arrow_factor, y = 0, yend = NMDS2* arrow_factor),
             arrow = arrow(length = unit(0.2, "cm")), colour = "#888888", alpha = 0.7) +
  geom_text_repel(data = spp.scrs.01, aes(x = NMDS1 * arrow_factor, y = NMDS2 * arrow_factor, 
                                          label = Species), size = 3, max.overlaps = getOption("ggrepel.max.overlaps", default = 50) ) +
  annotate("text", x = -.90, y = 1.1, label = "bold(Stress) == 0.177",
  parse = TRUE)+
  annotate("text", x = -1.05, y = 1, label = "bold(K) == 2",
  parse = TRUE) +
  theme_classic() +
  theme(text=element_text(size=16), legend.key.size=unit(0.5, "cm"))



NMDS_plot
```

**Figure 3.** Non-metric multidimensional scaling analysis (NMDS) ordination using Bray-Curtis dissimilarity visualizing seed rain species composition at a remnant prairie and prairie reconstructions of various ages (k = 2, stress = 0.174). Ellipses represent 95% confidence intervals calculated from standard error. Plot vectors represent species significantly associated with seed rain composition (p ≤ 0.001).

### Analysis 3: PERMANOVA

```{r, message = FALSE, echo = FALSE, warning = FALSE}


# Standardize using site total and bray-curtis
data.dist.tot <- vegdist(seed_mat_tot, method = "bray")
seed_com <- data.frame(site, seed_mat)
```

```{r, message = FALSE, echo = FALSE, warning = FALSE}


# PERMANOVA looking at differences in species composition between sites
seed_perma_mod_tot <- adonis2(data.dist.tot~as.factor(site), data=seed_com)
seed_perma_mod_tot

```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

# Pairwise comparisons between sites
pairwise.adonis2(data.dist.tot~as.factor(site), data=seed_com, p.adjust = "bonferroni")
```

### Analysis 3: Multivariate homogeneity of group dispersions test

```{r, message = FALSE, echo = FALSE, warning = FALSE}

# Betadisper looking at whether there are significant differences in community variation between sites.

## No significant differences

bd <- betadisper(data.dist.tot, as.factor(seed_com$site))
anova(bd)
```

### Analysis 3: Results

Species composition of the seed rain significantly differed between all sites (*PERMANOVA,* R^2^ = 0.41, F~3~ = 8.19, p \< 0.001; *all pairwise comparisons,* p \< 0.01). However, variation in community composition did not differ among the sampled prairies (*ANOVA,* R^2^ = 0.04,F~3~ = 0.54, p = 0.66). Reconstructed prairies of comparable ages shared the most similarities in regards to species composition. Except for some marginal overlap with the oldest reconstruction, the remnant prairie contained a distinct community of dispersing plants compared to reconstructed sites. In general, NMDS1 appeared to correspond to plant life history traits, where influential annual/biannual species in the seed rain align on the left side of the axis and perennial species on the right side. NMDS2 may be associated with plant functional group, where influential forbs cluster at higher values and graminoids at lower values along the axis.

## Analysis 4: Are there differences in the diversity of seed rain between sites?

### **Are there differences in the diversity of seed rain between sites?**

\-\--


```{r, results = 'hide', echo = FALSE, message = FALSE, warning = FALSE}

### Analysis 4: Data wrangling 

# Make a new dataset that aggregates the total number of seeds per species and per site for each week

Lumped_Seeds_SPP <- Seed_Data_Date %>% 
                        filter(!grepl('UNK',SPP6)) %>% 
                        filter(!grepl('PANICUM?',SPP6)) %>% 
                        group_by(Plot, Transect, SPP6) %>% 
                        summarize(totseeds = sum(Number))

Lumped_Seeds_SPP_date <- Seed_Data_Date %>% 
                        filter(!grepl('UNK',SPP6)) %>% 
                        filter(!grepl('PANICUM?',SPP6)) %>% 
                        group_by(Plot, Transect, SPP6, `Sampling Date`) %>% 
                        summarize(totseeds = sum(Number))


# For looking at total species richness
Seed_Data_Lumped <- full_join(Lumped_Seeds_SPP, Species_data)

Lumped_Seeds_SPP_date <- full_join(Lumped_Seeds_SPP_date , Species_data)

# Count the total number of distinct species per site according to unique transect 

Uni_Seed_Diversity <- Seed_Data_Lumped %>% 
  unite("UniqueTran", Plot:Transect, sep = "_", remove = FALSE) %>% 
  group_by(Plot, Transect, UniqueTran)  %>% 
   summarize(Species_richness = n_distinct(SPP6))

# Count the  number of distinct native species per site according to unique transect 

Native_Uni_Seed_Diversity <- Seed_Data_Lumped %>% 
  filter(Provenance == "Native") %>% 
  unite("UniqueTran", Plot:Transect, sep = "_", remove = FALSE) %>% 
  group_by(Plot, Transect, UniqueTran)  %>% 
   summarize(n_Species_richness = n_distinct(SPP6))


# Count the total number of distinct species per site according to sampling date
week_species_counts <- Lumped_Seeds_SPP_date  %>% 
                  filter(!grepl('UNK',SPP6)) %>% 
                  filter(!grepl('PANICUM?',SPP6)) %>% 
                  group_by(Plot, `Sampling Date`) %>% 
                  summarize(Species_richness = n_distinct(SPP6))

# Count the number of distinct native species per site according to sampling date

native_week_species_counts <- Lumped_Seeds_SPP_date   %>% 
                  filter(Provenance == "Native") %>% 
                  group_by(Plot, `Sampling Date`) %>% 
                  summarize(N_Species_richness = n_distinct(SPP6))
```



### Analysis 4: Fitting the models

#### Total Species Richness

```{r, message = FALSE, echo = FALSE, warning = FALSE}


# Make Tucker Prairie the reference level
Uni_Seed_Diversity$Plot <- factor(Uni_Seed_Diversity$Plot, levels=c("TP", "PFCA 1", "PFCA 2", "PFCA 3"))
str(Uni_Seed_Diversity$Plot)



# Model fitting

# -------

## Poisson 
### AIC = 256.68

## Negative binomial threw up an error message 
### AIC 258.68 + warning message

## Linear regression 
### AIC 258.65

# -------

# Decided to use the Poisson

Poisson_sr <- glm(Species_richness ~ Plot, family = "poisson", data = Uni_Seed_Diversity)
summary(Poisson_sr)

```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

## Do restored sites differ from the remnant?

# Contrasts
standardize <- emmeans(Poisson_sr, "Plot")
contrast(standardize, "trt.vs.ctrl")


## 95% confidence intervals 

emmeans::emmeans(Poisson_sr, "Plot", type = "response")

## Make a new dataframe to hold the model estimates and 95% confidence interval bounds

response.sr <- c(29.5, 35.8, 40.9, 27.2)
lower.sr <- c(26.3, 32.3, 37.1, 24.2)
upper.sr <- c(33.1, 39.7, 45.1, 30.6)
site <- c("TP", "PFCA 1", "PFCA 2", "PFCA 3")

Poisson.sr.est <- data.frame(site, response.sr ,lower.sr, upper.sr)

```

#### Native Species Richness

```{r, message = FALSE, echo = FALSE, warning = FALSE}


# Make Tucker Prairie the reference level
Native_Uni_Seed_Diversity$Plot <- factor(Native_Uni_Seed_Diversity$Plot, levels=c("TP", "PFCA 1", "PFCA 2", "PFCA 3"))
str(Native_Uni_Seed_Diversity$Plot)



# Model fitting

# -------

## Poisson 
### AIC = 246.46

## Negative binomial threw up an error message 
### AIC 248.46 + warning message

## Linear regression 
### AIC 246.82

# -------

# Decided to use the Poisson 

Poisson_sr_native <- glm(n_Species_richness ~ Plot, family = "poisson", data = Native_Uni_Seed_Diversity)
summary(Poisson_sr_native )
```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

## Do restored sites differ from the remnant?

# Contrasts
standardize <- emmeans(Poisson_sr_native, "Plot")
contrast(standardize, "trt.vs.ctrl")

## 95% confidence intervals 

emmeans::emmeans(Poisson_sr_native, "Plot", type = "response")

## Make a new dataframe to hold the model estimates and 95% confidence interval bounds

response.sr.native <- c(27.4, 28.1, 33.0, 24.1)
lower.sr.native <- c(24.3, 25.0, 29.6, 21.2)
upper.sr.native <- c(30.8, 31.6, 36.8, 27.3)
site <- c("TP", "PFCA 1", "PFCA 2", "PFCA 3")

Poisson.sr.est.native <- data.frame(site, response.sr.native, lower.sr.native, upper.sr.native)

```

#### Shannon Diversity

```{r, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE}

# Shannon's Diversity

shannon_com_mat <- spread(Lumped_Seeds_SPP, SPP6, totseeds, fill = 0)

#Calculate Shannon's Diversity

Shannon <- diversity(shannon_com_mat[,-c(1,2)], index="shannon")

# Make it into a dataframe
Shannon <- as.data.frame(Shannon)

# Add back the column with the site labels
Shannon$Site <- seed_com[,1]

favstats(Shannon$Shannon~Shannon$Site)

```

```{r}

# Make Tucker Prairie the reference level

Shannon$Site <- factor(Shannon$Site, levels=c("TP", "PFCA 1", "PFCA 2", "PFCA 3"))

Shannon_sr <- lm(Shannon ~ Site, data = Shannon)
summary(Shannon_sr)
anova(Shannon_sr)


standardize <- emmeans(Shannon_sr, "Site")
contrast(standardize, "trt.vs.ctrl")


```




### Analysis 4: Data visualization

#### Species Richness

```{r, message = FALSE, echo = FALSE, warning = FALSE}


Uni_Seed_Diversity$Plot <- factor(Uni_Seed$Plot, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))


Poisson.sr.est$site <- factor(Poisson.sr.est$site, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))

SR.plot <- ggplot(data=Uni_Seed_Diversity)+
  geom_jitter(data=Uni_Seed_Diversity, aes(x=Plot, y=Species_richness, color = Plot), show.legend = FALSE, width = .2, size = 2) +
  geom_point(data = Poisson.sr.est, aes(x=site, y=response.sr), show.legend = FALSE, size = 3.5)+
  geom_errorbar(data = Poisson.sr.est, aes(x= site, y = response.sr, ymin = lower.sr, ymax = upper.sr), width = 0.15, cex = 0.75)+
  theme_classic()+
  xlab("")+
  ylim(0,60)+
  ylab("Total richness")+
  scale_x_discrete(labels = c ("PFCA 1" = "Young", "PFCA 2" = "Middle", "PFCA 3" = "Old", "TP" = "Remnant"), guide = guide_axis(angle = 0)) +
  scale_color_manual(name = "Site", values=c( "#1f78b4", "#a6cee3", "#b2df8a","#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant"))+
 theme(text=element_text(size=14), legend.key.size=unit(1, "cm"))

SR.plot.annotated <- SR.plot +  annotate("text", x = c(1,2) , y = 55, label = "*", size = 9.5)+
  geom_vline(xintercept = 3.5, linetype = "longdash")


```



```{r, message = FALSE, echo = FALSE, warning = FALSE}

Native_Uni_Seed_Diversity$Plot <- factor(Native_Uni_Seed_Diversity$Plot, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))


Poisson.sr.est.native$site <- factor(Poisson.sr.est.native$site, levels=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"))

Native.SR.plot.annotated <- ggplot()+
  geom_jitter(data=Native_Uni_Seed_Diversity, aes(x=Plot, y=n_Species_richness, color = Plot), show.legend = FALSE, width = .2, size = 2) +
  geom_point(data = Poisson.sr.est.native, aes(x=site, y=response.sr.native), show.legend = FALSE, size = 3.5)+
  geom_errorbar(data = Poisson.sr.est.native, aes(x= site, y = response.sr.native, ymin = lower.sr.native, ymax = upper.sr.native), width = 0.15, cex = 0.75)+
  theme_classic()+
  xlab("")+
  ylim(0,60)+
  ylab("Native richness")+
  scale_x_discrete(labels = c ("PFCA 1" = "Young", "PFCA 2" = "Middle", "PFCA 3" = "Old", "TP" = "Remnant"), guide = guide_axis(angle = 0)) +
  scale_color_manual(name = "Site", values=c( "#1f78b4", "#a6cee3", "#b2df8a","#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant"))+
 theme(text=element_text(size=14), legend.key.size=unit(1, "cm"))


Native.SR.plot.annotated <- Native.SR.plot.annotated + geom_vline(xintercept = 3.5, linetype = "longdash")


```



```{r, message = FALSE, echo = FALSE, warning = FALSE}


# setwd("~/Desktop/Seed Rain 2019/Graphics")
Tot_SR_fig <- ggplot(week_species_counts, aes(y=Species_richness, x=`Sampling Date`, group=Plot)) +
  geom_line(aes(colour=Plot), size = 1.5,  show.legend = FALSE) +
  geom_point(aes(shape=Plot, fill=Plot), size=3,  show.legend = FALSE)+
  scale_color_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019"))+
    scale_fill_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019"))+
  scale_shape_manual(name = "", values=c(23,24,22,21),breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant 2019")) +
  theme_classic()+
  ylim(0,60)+
  scale_x_datetime(date_breaks = '1 month', date_labels="%b %d", limits=as.POSIXct(c("2019-06-10", "2020-01-01")))+
  theme(text=element_text(size=14), legend.key.size=unit(1, "cm"))+
  ylab("Total richness")+
  xlab("")+
  guides(x = guide_axis(angle = 45))

```



```{r, message = FALSE, echo = FALSE, warning = FALSE}


# setwd("~/Desktop/Seed Rain 2019/Graphics")
Tot_N_SR_fig <- ggplot(native_week_species_counts, aes(y=N_Species_richness, x=`Sampling Date`, group=Plot)) +
  geom_line(aes(colour=Plot), size = 1.5) +
  geom_point(aes(shape=Plot, fill=Plot), size=3)+
  scale_color_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant"))+
    scale_fill_manual(name = "", values=c("#1f78b4", "#a6cee3", "#b2df8a", "#33a02c"), breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant"))+
  scale_shape_manual(name = "", values=c(23,24,22,21),breaks=c("PFCA 1", "PFCA 2", "PFCA 3", "TP"), labels = c("Young", "Middle", "Old", "Remnant")) +
  theme_classic()+
  ylim(0,60)+
  scale_x_datetime(date_breaks = '1 month', date_labels="%b %d", limits=as.POSIXct(c("2019-06-10", "2020-01-01")))+
  theme(text=element_text(size=14), legend.key.size=unit(0.5, "cm"))+
  ylab("Native richness")+
  xlab("")+
  guides(x = guide_axis(angle = 45))

```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

legend <- get_legend(Tot_N_SR_fig)
```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

panel_1 <- plot_grid(SR.plot.annotated, Native.SR.plot.annotated,Tot_SR_fig,Tot_N_SR_fig + theme(legend.position = 'none'),
          labels=c("A","B","C","D", ncol = 2),
          rel_heights = c(0.9,1))

plot_grid(panel_1 , legend, rel_widths = c(3,.5), rel_heights = c(1,2))

```


**Results :** Mean total morphospecies richness was only significantly greater in the youngest and middle-aged prairie compared to the remnant. While only mean native morphospecies richness was greater in the middle-aged prairie. Overall, morphospecies richness trended upward during the study period, with more morphospecies captured in the late fall than in the summer.



## Analysis 5: Supplemental table to assist in cross study comparison

```{r, message = FALSE, echo = FALSE, warning = FALSE}

Lumped_unknown1 <- Seed_Data_Date  %>% 
  filter(grepl('UNK',SPP6)) %>% 
  group_by(Plot, `Sampling Date`) %>% 
                        summarize(totseeds = sum(Number))

Lumped_unknown2 <- Seed_Data_Date %>% 
  filter(grepl('PANICUM?',SPP6)) %>% 
  group_by(Plot, `Sampling Date`) %>% 
                        summarize(totseeds = sum(Number))

Lumped_unknown <- full_join(Lumped_unknown1, Lumped_unknown2)

Lumped_unknown$SPP6 <- rep("UNKSPP", 24)

Lumped_no_unknown <- Seed_Data_Date %>% 
   filter(!grepl('UNK',SPP6)) 

Lumped_no_unknown <- Lumped_no_unknown  %>% 
   filter(!grepl('PANICUM?',SPP6)) %>%
    group_by(Plot, `Sampling Date`, SPP6) %>% 
                        summarize(totseeds = sum(Number))

full_SPP6_Date <- full_join(Lumped_unknown, Lumped_no_unknown)


full_SPP6_Date_TOT <- full_SPP6_Date %>% 
                        group_by(SPP6, `Sampling Date`) %>% 
                        summarize(totseeds = sum(totseeds))

```

```{r}
Lumped_full_SPP6 <- full_SPP6_Date %>%
                         group_by(Plot, SPP6) %>% 
                         summarize(totseeds = sum(totseeds))
  
Lumped_full_SPP6_nu <- Lumped_full_SPP6 %>% 
   filter(!grepl('UNK',SPP6)) %>% 
   summarize(totseeds = sum(totseeds))
   
Lumped_full_SPP6_u <- Lumped_full_SPP6 %>% 
   filter(grepl('UNK',SPP6))


Lumped_full_SPP6_nu$percent_known <- 100-(Lumped_full_SPP6_u$totseeds / Lumped_full_SPP6_nu$totseeds)*100

Lumped_full_SPP6_nu
```



```{r, message = FALSE, echo = FALSE, warning = FALSE}

# Per site
seed_community_phenology_site <- spread(full_SPP6_Date, `Sampling Date`, totseeds, fill = 0)

# Lumped across site

seed_community_phenology_lumped <- spread(full_SPP6_Date_TOT, `Sampling Date`, totseeds, fill = 0)


seed_community_phenology_lumped <- seed_community_phenology_lumped %>% 
  arrange(desc(`2019-06-13`), desc(`2019-06-27`), desc(`2019-07-10`), desc(`2019-07-11`), desc(`2019-07-26`), desc(`2019-08-08`),  desc(`2019-08-22`),  desc(`2019-09-05`), desc(`2019-09-19`), desc(`2019-10-03`), desc(`2019-10-17`), desc(`2019-11-02`), desc(`2019-11-14`), desc(`2019-12-01`), desc(`2019-12-12`))


seed_community_phenology_lumped [seed_community_phenology_lumped  == 0] <- NA
```

```{r}

A1 <- seed_community_phenology_lumped %>% mutate_if(is.numeric, str_replace_all, pattern = "NA", replacement = ".")

ft <- flextable(A1)

ft

```

```{r, message = FALSE, echo = FALSE, warning = FALSE}

#read_docx() %>% 
  #body_add_flextable(value = ft) %>%  
  #print(target = "example_table_word.docx")



```


```{r, message = FALSE, echo = FALSE, warning = FALSE}
# Species lists for each site

### How many species at each site?

all_species <- Seed_Data_Lumped %>% 
  group_by(Plot)  %>% 
   summarize(Species_richness = n_distinct(SPP6))

all_species

### Species found at TP

TP_species <- Seed_Data_Lumped %>% 
  filter(Plot == "TP") %>% 
  group_by(Plot, SPP6)  %>% 
   summarize(tot = sum(totseeds))


### Species found at PFCA 1

PFCA1_species <- Seed_Data_Lumped %>% 
  filter(Plot == "PFCA 1") %>% 
  group_by(Plot, SPP6)  %>% 
   summarize(tot = sum(totseeds))

### Species found at PFCA 2

PFCA2_species <- Seed_Data_Lumped %>% 
  filter(Plot == "PFCA 2") %>% 
  group_by(Plot, SPP6)  %>% 
   summarize(tot = sum(totseeds))


### Species found at PFCA 3


PFCA3_species <- Seed_Data_Lumped %>% 
  filter(Plot == "PFCA 3") %>% 
  group_by(Plot, SPP6)  %>% 
   summarize(tot = sum(totseeds))
```


```{r}

SPP6_Frequency <- full_SPP6_Date %>% 
                        group_by(SPP6,) %>% 
                        summarize(totseeds = sum(totseeds))

```
